The curious world of constructive mathematics

### 2022-09-03 algar-lecture-notes.tex
typo
to do

\begin{exercise}[Constructive status of classical tautologies]
Which of the following classical tautologies can reasonably be expected to
admit constructive proofs?
\begin{enumerate}
\item $\neg(\alpha \vee \beta) \ \Longrightarrow\ \neg\alpha \wedge \neg\beta$
\item $\neg(\alpha \wedge \beta) \ \Longrightarrow\ \neg\alpha \vee \neg\beta$
\item $(\alpha \vee \beta) \wedge \neg\alpha \ \Longrightarrow\ \beta$
\item $\forall M \in P(X)\_ (\exists x\?X\_ x \in M) \vee M = \emptyset$
(already interesting for~$X = \{\star\}$)
\item $\forall n\?\NN\_ (n = 0 \vee n \neq 0)$
\item $\forall x\?\RR\_ (x = 0 \vee x \neq 0)$
\item $\forall x\?\RR\_ (\neg(\exists y\?\RR\_ xy=1) \Rightarrow x = 0)$
\item $\forall z\?\overline{\QQ}\_ (z = 0 \vee z \neq 0)$
\item $\forall f \? \NN \to \{0,1\}\_ (\neg\neg\exists n \? \NN\_ f(n) = 0)
\Rightarrow (\exists n \? \NN\_ f(n) = 0)$
\item $\forall f \? \NN \to \{0,1\}\_ \exists n \? \NN\_ (f(n) = 1
\Leftrightarrow (\forall m \? \NN\_ f(m) = 1))$
\item $\forall f \? \NN \to \{0,1\}\_ (\exists n \? \NN\_ f(n) = 0) \vee
(\forall n \? \NN\_ f(n) = 1)$
\item $\forall f \? \NN_\infty \to \{0,1\}\_ (\exists n \? \NN_\infty\_ f(n) = 0) \vee
(\forall n \? \NN_\infty\_ f(n) = 1)$

{\noindent\scriptsize\emph{Remark.} The set~$\NN_\infty$ is the \emph{one-point
compactification} of~$\NN$. A sensible definition in constructive mathematics
is as the set of decreasing binary sequences~$(x_0,x_1,x_2,\ldots)$. The
naturals embed into~$\NN_\infty$ by mapping~$n$ to the sequence~$1^n 0^\omega =
(1,\ldots,1,0,\ldots,0)$, and an element not in the image of this embedding
is~$\infty \defeq 1^\omega = (1,1,\ldots)$. Assuming the law of excluded
middle (or already weaker principles), every element of~$\NN_\infty$ is of one
of these two forms. Martín Escardó has worked extensively on unexpected
instances of the principle of omniscience for searchable sets
like~$\NN_\infty$~\cite{escardo:omniscience1,escardo:omniscience2,escardo:omniscience}.\par}
\end{enumerate}
\end{exercise}

@@

>>>>

\begin{exercise}[Constructive status of classical tautologies]
Which of the following classical tautologies can reasonably be expected to
admit constructive proofs?
\begin{enumerate}
\item $\neg(\alpha \vee \beta) \ \Longrightarrow\ \neg\alpha \wedge \neg\beta$
\item $\neg(\alpha \wedge \beta) \ \Longrightarrow\ \neg\alpha \vee \neg\beta$
\item $(\alpha \vee \beta) \wedge \neg\alpha \ \Longrightarrow\ \beta$
\item $\forall M\?P(X)\_ (\exists x\?X\_ x \in M) \vee M = \emptyset$
(already interesting for~$X = \{\star\}$)
\item $\forall n\?\NN\_ (n = 0 \vee n \neq 0)$
\item $\forall x\?\RR\_ (x = 0 \vee x \neq 0)$
\item $\forall x\?\RR\_ (\neg(\exists y\?\RR\_ xy=1) \Rightarrow x = 0)$
\item $\forall z\?\overline{\QQ}\_ (z = 0 \vee z \neq 0)$
\item $\forall f \? \NN \to \{0,1\}\_ (\neg\neg\exists n \? \NN\_ f(n) = 0)
\Rightarrow (\exists n \? \NN\_ f(n) = 0)$ (Markov's principle)
\item $\forall f \? \NN \to \{0,1\}\_ \exists n \? \NN\_ (f(n) = 1
\Rightarrow (\forall m \? \NN\_ f(m) = 1))$ (Drinker's paradox)
\item $\forall f \? \NN \to \{0,1\}\_ (\exists n \? \NN\_ f(n) = 0) \vee
(\forall n \? \NN\_ f(n) = 1)$
\item $\forall f \? \NN_\infty \to \{0,1\}\_ (\exists n \? \NN_\infty\_ f(n) = 0) \vee
(\forall n \? \NN_\infty\_ f(n) = 1)$

{\noindent\scriptsize\emph{Remark.} The set~$\NN_\infty$ is the \emph{one-point
compactification} of~$\NN$. A sensible definition of it in constructive mathematics
is as the set of decreasing binary sequences~$(x_0,x_1,x_2,\ldots)$. The
naturals embed into~$\NN_\infty$ by mapping~$n$ to the sequence~$1^n 0^\omega =
(1,\ldots,1,0,\ldots,0)$, and an element not in the image of this embedding
is~$\infty \defeq 1^\omega = (1,1,\ldots)$. Assuming the law of excluded
middle (or already weaker principles), every element of~$\NN_\infty$ is of one
of these two forms. Martín Escardó has worked extensively on unexpected
instances of the principle of omniscience for searchable sets
like~$\NN_\infty$~\cite{escardo:omniscience1,escardo:omniscience2,escardo:omniscience}.\par}
\end{enumerate}
\end{exercise}

@@

1
Add Agda exercises

2
classical logic fairy tale
Add Agda exercises
Mention Dialectica
something with minima (for negneg translation)
Exercises: negneg proof, stability of the HA axioms, applications of Barr
negneg drinker and CPS perspective

3
Definition
Examples
Language: definition, properties, IVT, FTA
Exercises: dense, monotone IVT, ...

4
"A fantastical ring"
The generic prime filter
Explicit construction
Exercises: (lift from Spiel & Spaß)
